---
title: "covid_manuscript"
author: "Sucharitha Dodamgodage"
date: "2023-02-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Packages
```{r}
library(GGally)
library(dplyr)#combine datasets
library(plyr)
library(ggplot2)
library(Hmisc) #correlation with p value
library(knitr)
library(tidyverse,warn.conflicts = F)
library(tidyr)
library(tibble)
library(purrr)
library(sp)
library(moments)
library(MASS)
library(foreign)
library(car)
library(readxl)
library(FSA)
library(ggpubr)#ggarrange
library(svglite)
library(formattable)
library(VGAM)#Generalised poisson
library(gamlss)
library(gamlss.dist)
library(cluster)
library(factoextra)
library(Metrics)#RMSE
library(Ckmeans.1d.dp)
library(readxl)
```

Insert data
```{r}
dataphase1<- read_xlsx("Data/Phase1_complete.xlsx")
dataphase2 <- read_xlsx("Data/Phase2_complete.xlsx")
datafull <- read_xlsx("Data/All_year_complete.xlsx")
```
Split the all year data set based on infection rate cluster
```{r}
cluster_low_inf_datafull<- subset(datafull,datafull$Infection_rate_clusters=="Low")
cluster_medium_inf_datafull<- subset(datafull,datafull$Infection_rate_clusters=="Medium")
cluster_high_inf_datafull<- subset(datafull,datafull$Infection_rate_clusters=="High")
```

Split the all year data set based on fatality rate cluster
```{r}
cluster_low_fat_datafull<- subset(datafull,datafull$Fatality_rate_clusters=="Low")
cluster_medium_fat_datafull<- subset(datafull,datafull$Fatality_rate_clusters=="Medium")
cluster_high_fat_datafull<- subset(datafull,datafull$Fatality_rate_clusters=="High")
```
Split the phase 1 data set based on infection rate cluster
```{r}
cluster_low_inf_dataphase1<- subset(dataphase1,dataphase1$Infection_rate_clusters=="Low")
cluster_medium_inf_dataphase1<- subset(dataphase1,dataphase1$Infection_rate_clusters=="Medium")
cluster_high_inf_dataphase1<- subset(dataphase1,dataphase1$Infection_rate_clusters=="High")
```

Split the phase 1 data set based on fatality rate cluster
```{r}
cluster_low_fat_dataphase1<- subset(dataphase1,dataphase1$Fatality_rate_clusters=="Low")
cluster_medium_fat_dataphase1<- subset(dataphase1,dataphase1$Fatality_rate_clusters=="Medium")
cluster_high_fat_dataphase1<- subset(dataphase1,dataphase1$Fatality_rate_clusters=="High")
```

Split the phase 2 data set based on infection rate cluster
```{r}
cluster_low_inf_dataphase2<- subset(dataphase2,dataphase2$Infection_rate_clusters=="Low")
cluster_medium_inf_dataphase2<- subset(dataphase2,dataphase2$Infection_rate_clusters=="Medium")
cluster_high_inf_dataphase2<- subset(dataphase2,dataphase2$Infection_rate_clusters=="High")
```

Split the phase 2 data set based on fatality rate cluster
```{r}
cluster_low_fat_dataphase2<- subset(dataphase2,dataphase2$Fatality_rate_clusters=="Low")
cluster_medium_fat_dataphase2<- subset(dataphase2,dataphase2$Fatality_rate_clusters=="Medium")
cluster_high_fat_dataphase2<- subset(dataphase2,dataphase2$Fatality_rate_clusters=="High")
```


#############################################
Mean,Variance,Skewness and Kurtosis
#############################################

#All year_low_infection
```{r}
mean(cluster_low_inf_datafull$infection_rate_10k)
sd(cluster_low_inf_datafull$infection_rate_10k)
kurtosis(cluster_low_inf_datafull$infection_rate_10k)
skewness(cluster_low_inf_datafull$infection_rate_10k)
```

#All year_medium_infection
```{r}
mean(cluster_medium_inf_datafull$infection_rate_10k)
sd(cluster_medium_inf_datafull$infection_rate_10k)
kurtosis(cluster_medium_inf_datafull$infection_rate_10k)
skewness(cluster_medium_inf_datafull$infection_rate_10k)
```

#All year_high_infection
```{r}
mean(cluster_high_inf_datafull$infection_rate_10k)
sd(cluster_high_inf_datafull$infection_rate_10k)
kurtosis(cluster_high_inf_datafull$infection_rate_10k)
skewness(cluster_high_inf_datafull$infection_rate_10k)
```

#phase1_low_infection

```{r}
mean(cluster_low_inf_dataphase1$infection_rate_10k)
sd(cluster_low_inf_dataphase1$infection_rate_10k)
kurtosis(cluster_low_inf_dataphase1$infection_rate_10k)
skewness(cluster_low_inf_dataphase1$infection_rate_10k)
```

#phase1_medium_infection

```{r}
mean(cluster_medium_inf_dataphase1$infection_rate_10k)
sd(cluster_medium_inf_dataphase1$infection_rate_10k)
kurtosis(cluster_medium_inf_dataphase1$infection_rate_10k)
skewness(cluster_medium_inf_dataphase1$infection_rate_10k)
```

#phase1_high_infection

```{r}
mean(cluster_high_inf_dataphase1$infection_rate_10k)
sd(cluster_high_inf_dataphase1$infection_rate_10k)
kurtosis(cluster_high_inf_dataphase1$infection_rate_10k)
skewness(cluster_high_inf_dataphase1$infection_rate_10k)
```

#phase2_low_infection
```{r}
mean(cluster_low_inf_dataphase2$infection_rate_10k)
sd(cluster_low_inf_dataphase2$infection_rate_10k)
kurtosis(cluster_low_inf_dataphase2$infection_rate_10k)
skewness(cluster_low_inf_dataphase2$infection_rate_10k)
```

#phase2_medium_infection
```{r}
mean(cluster_medium_inf_dataphase2$infection_rate_10k)
sd(cluster_medium_inf_dataphase2$infection_rate_10k)
kurtosis(cluster_medium_inf_dataphase2$infection_rate_10k)
skewness(cluster_medium_inf_dataphase2$infection_rate_10k)
```
#phase2_high_infection
```{r}
mean(cluster_high_inf_dataphase2$infection_rate_10k)
sd(cluster_high_inf_dataphase2$infection_rate_10k)
kurtosis(cluster_high_inf_dataphase2$infection_rate_10k)
skewness(cluster_high_inf_dataphase2$infection_rate_10k)
```

#all year_low_fatality
```{r}
mean(cluster_low_fat_datafull$infection_rate_10k)
sd(cluster_low_fat_datafull$infection_rate_10k)
kurtosis(cluster_low_fat_datafull$infection_rate_10k)
skewness(cluster_low_fat_datafull$infection_rate_10k)
```
#all year_medium_fatality
```{r}
mean(cluster_medium_fat_datafull$infection_rate_10k)
sd(cluster_medium_fat_datafull$infection_rate_10k)
kurtosis(cluster_medium_fat_datafull$infection_rate_10k)
skewness(cluster_medium_fat_datafull$infection_rate_10k)
```
#all year_high_fatality
```{r}
mean(cluster_high_fat_datafull$infection_rate_10k)
sd(cluster_high_fat_datafull$infection_rate_10k)
kurtosis(cluster_high_fat_datafull$infection_rate_10k)
skewness(cluster_high_fat_datafull$infection_rate_10k)
```
#phase1_low_fatality
```{r}
mean(cluster_low_fat_dataphase1$infection_rate_10k)
sd(cluster_low_fat_dataphase1$infection_rate_10k)
kurtosis(cluster_low_fat_dataphase1$infection_rate_10k)
skewness(cluster_low_fat_dataphase1$infection_rate_10k)
```
#phase1_medium_fatality
```{r}
mean(cluster_medium_fat_dataphase1$infection_rate_10k)
sd(cluster_medium_fat_dataphase1$infection_rate_10k)
kurtosis(cluster_medium_fat_dataphase1$infection_rate_10k)
skewness(cluster_medium_fat_dataphase1$infection_rate_10k)
```
#phase1_high_fatality
```{r}
mean(cluster_high_fat_dataphase1$infection_rate_10k)
sd(cluster_high_fat_dataphase1$infection_rate_10k)
kurtosis(cluster_high_fat_dataphase1$infection_rate_10k)
skewness(cluster_high_fat_dataphase1$infection_rate_10k)
```
#phase2_low_fatality
```{r}
mean(cluster_low_fat_dataphase2$infection_rate_10k)
sd(cluster_low_fat_dataphase2$infection_rate_10k)
kurtosis(cluster_low_fat_dataphase2$infection_rate_10k)
skewness(cluster_low_fat_dataphase2$infection_rate_10k)
```
#phase2_medium_fatality
```{r}
mean(cluster_medium_fat_dataphase2$infection_rate_10k)
sd(cluster_medium_fat_dataphase2$infection_rate_10k)
kurtosis(cluster_medium_fat_dataphase2$infection_rate_10k)
skewness(cluster_medium_fat_dataphase2$infection_rate_10k)
```
#phase2_high_fatality
```{r}
mean(cluster_high_fat_dataphase2$infection_rate_10k)
sd(cluster_high_fat_dataphase2$infection_rate_10k)
kurtosis(cluster_high_fat_dataphase2$infection_rate_10k)
skewness(cluster_high_fat_dataphase2$infection_rate_10k)
```

#############################
KW test
###########################
```{r}
#####infection
#All year 
#normality check
shapiro.test(datafull$popdense)

#KW test
kruskal.test( datafull$pct_Adults_with_Obesity ~ datafull$Infection_rate_clusters, datafull)

dunnTest(datafull$pct_Adults_with_Obesity~ datafull$Infection_rate_clusters, datafull,two.sided = TRUE,method = "none")

#Phase 1
#normality check
shapiro.test(dataphase1$popdense)

#KW test
kruskal.test( dataphase1$pct_Adults_with_Diabetes ~ dataphase1$Infection_rate_clusters, dataphase1)

dunnTest(dataphase1$pct_Adults_with_Diabetes~ dataphase1$Infection_rate_clusters, dataphase1,two.sided = TRUE,method = "none")

#Phase 2
#normality check
shapiro.test(dataphase2$popdense)

#KW test
kruskal.test( dataphase2$pct_Adults_with_Obesity ~ dataphase2$Infection_rate_clusters, dataphase2)

dunnTest(dataphase2$pct_Adults_with_Obesity~ dataphase2$Infection_rate_clusters, dataphase2,two.sided = TRUE,method = "none")


#####fatality
#All year 
#normality check
shapiro.test(datafull$popdense)

#KW test
kruskal.test( datafull$pct_Adults_with_Obesity ~ datafull$Fatality_rate_clusters, datafull)

dunnTest(datafull$pct_Adults_with_Obesity~ datafull$Fatality_rate_clusters, datafull,two.sided = TRUE,method = "none")

#Phase 1
#normality check
shapiro.test(dataphase1$popdense)

#KW test
kruskal.test( dataphase1$pct_Adults_with_Obesity ~ dataphase1$Fatality_rate_clusters, dataphase1)

dunnTest(dataphase1$pct_Adults_with_Obesity~ dataphase1$Fatality_rate_clusters, dataphase1,two.sided = TRUE,method = "none")

#Phase 2
#normality check
shapiro.test(dataphase2$popdense)

#KW test
kruskal.test( dataphase2$pct_Adults_with_Obesity ~ dataphase2$Fatality_rate_clusters, dataphase2)

dunnTest(dataphase2$pct_Adults_with_Obesity~ dataphase2$Fatality_rate_clusters, dataphase2,two.sided = TRUE,method = "none")


############################
Regression Analysis
############################

INFECTION

#Phase 1 High infection model 1
```{r}
#Removing outliers
summary(cluster_high_inf_dataphase1$infection_rate_10k)
iqr<-IQR(cluster_high_inf_dataphase1$infection_rate_10k)
Tmax = 149.20+3*iqr #consider extreme outlires
Tmin = 80.05-3*iqr #consider extreme outlires
#outliers
outl<- subset(cluster_high_inf_dataphase1,cluster_high_inf_dataphase1$infection_rate_10k<Tmin| cluster_high_inf_dataphase1$infection_rate_10k>Tmax)
#cleaned data
cleaned_data<- subset(cluster_high_inf_dataphase1,cluster_high_inf_dataphase1$infection_rate_10k>Tmin & cluster_high_inf_dataphase1$infection_rate_10k<Tmax)
#normality test
shapiro.test(cleaned_data$infection_rate_10k)

#summary(cleaned_data$infection_rate_10k)
#Normalize data(since data is not normally distributed we cannot use z score)
#cleaned_data$normalized_infection_rate_10k<- (cleaned_data$infection_rate_10k-63.17)/(321.49-63.17)

#(But fitting NB model to a standardized data doesn't make any sense.Therefore we decided to fit a model after removing outliers)

#Now I fit the NB model for infection_rate_10k in cleaned data

#convert infection rate into integers
cleaned_data$infection_rate_10k_integer<-round(cleaned_data$infection_rate_10k,digits = 0)

#split data set into training and testing
#make this example reproducible
set.seed(1)

#use 70% of dataset as training set and 30% as test set
sample <- sample(c(TRUE, FALSE), nrow(cleaned_data), replace=TRUE, prob=c(0.8,0.2))
train  <- cleaned_data[sample, ]
test   <- cleaned_data[!sample, ]

#fit NB model to training set
summary(m1 <- glm.nb(infection_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pctwhite+pctblack+pcthispanic+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity,data = train))
vif(m1)

#remove high vif(>5)
summary(m1 <- glm.nb(infection_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pctblack+pcthispanic+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity,data = train))
vif(m1)



#RMSE for train data
#create new data set using independent varaibles
train_predict<- subset(train, select = -c(infection_rate_10k_integer))
#predict usin fitted NB 
train$predict_infection<- predict(m1, train_predict,type ="response")
rmse_train<-rmse(train$infection_rate_10k_integer,train$predict_infection)
rmse_train

#Normalized RMSE train
Normalized_rmse_train<- rmse_train/mean(train$infection_rate_10k_integer)
Normalized_rmse_train


#RMSE for test data
test_predict<- subset(test, select = -c(infection_rate_10k_integer))
test$predict_infection<- predict(m1, test_predict,type = "response")
rmse_test<- rmse(test$infection_rate_10k_integer,test$predict_infection)
rmse_test


#Normalized RMSE test
Normalized_rmse_test<- rmse_test/mean(test$infection_rate_10k_integer)
Normalized_rmse_test
```

#Phase 1 Medium infection model 2
```{r}
#Removing outliers
summary(cluster_medium_inf_dataphase1$infection_rate_10k)
iqr<-IQR(cluster_medium_inf_dataphase1$infection_rate_10k)
Tmax = 42.09+3*iqr #consider extreme outlires
Tmin = 19.92-3*iqr #consider extreme outlires
#outliers
outl<- subset(cluster_medium_inf_dataphase1,cluster_medium_inf_dataphase1$infection_rate_10k<Tmin| cluster_medium_inf_dataphase1$infection_rate_10k>Tmax)
#cleaned data
cleaned_data<- subset(cluster_medium_inf_dataphase1,cluster_medium_inf_dataphase1$infection_rate_10k>Tmin & cluster_medium_inf_dataphase1$infection_rate_10k<Tmax)
#normality test
shapiro.test(cleaned_data$infection_rate_10k)

#summary(cleaned_data$infection_rate_10k)
#Normalize data(since data is not normally distributed we cannot use z score)
#cleaned_data$normalized_infection_rate_10k<- (cleaned_data$infection_rate_10k-63.17)/(321.49-63.17)

#(But fitting NB model to a standardized data doesn't make any sense.Therefore we decided to fit a model after removing outliers)

#Now I fit the NB model for infection_rate_10k in cleaned data

#convert infection rate into integers
cleaned_data$infection_rate_10k_integer<-round(cleaned_data$infection_rate_10k,digits = 0)

#split data set into training and testing
#make this example reproducible
set.seed(1)

#use 70% of dataset as training set and 30% as test set
sample <- sample(c(TRUE, FALSE), nrow(cleaned_data), replace=TRUE, prob=c(0.8,0.2))
train  <- cleaned_data[sample, ]
test   <- cleaned_data[!sample, ]

#fit NB model to training set
summary(m1 <- glm.nb(infection_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pctwhite+pctblack+pcthispanic+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity,data = train))
vif(m1)

#remove high vif(>5)
summary(m1 <- glm.nb(infection_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pctblack+pcthispanic+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity,data = train))
vif(m1)

#RMSE for train data
#create new data set using independent varaibles
train_predict<- subset(train, select = -c(infection_rate_10k_integer))
#predict usin fitted NB 
train$predict_infection<- predict(m1, train_predict,type ="response")
rmse_train<-rmse(train$infection_rate_10k_integer,train$predict_infection)
rmse_train

#Normalized RMSE train
Normalized_rmse_train<- rmse_train/mean(train$infection_rate_10k_integer)
Normalized_rmse_train


#RMSE for test data
test_predict<- subset(test, select = -c(infection_rate_10k_integer))
test$predict_infection<- predict(m1, test_predict,type = "response")
rmse_test<- rmse(test$infection_rate_10k_integer,test$predict_infection)
rmse_test


#Normalized RMSE test
Normalized_rmse_test<- rmse_test/mean(test$infection_rate_10k_integer)
Normalized_rmse_test
```

#Phase 1 Low infection model 3
```{r}
#Removing outliers
summary(cluster_low_inf_dataphase1$infection_rate_10k)
iqr<-IQR(cluster_low_inf_dataphase1$infection_rate_10k)
Tmax = 9.8723+3*iqr #consider extreme outlires
Tmin = 4.8041-3*iqr #consider extreme outlires
#outliers
outl<- subset(cluster_low_inf_dataphase1,cluster_low_inf_dataphase1$infection_rate_10k<Tmin| cluster_low_inf_dataphase1$infection_rate_10k>Tmax)
#cleaned data
cleaned_data<- subset(cluster_low_inf_dataphase1,cluster_low_inf_dataphase1$infection_rate_10k>Tmin & cluster_low_inf_dataphase1$infection_rate_10k<Tmax)
#normality test
shapiro.test(cleaned_data$infection_rate_10k)

#summary(cleaned_data$infection_rate_10k)
#Normalize data(since data is not normally distributed we cannot use z score)
#cleaned_data$normalized_infection_rate_10k<- (cleaned_data$infection_rate_10k-63.17)/(321.49-63.17)

#(But fitting NB model to a standardized data doesn't make any sense.Therefore we decided to fit a model after removing outliers)

#Now I fit the NB model for infection_rate_10k in cleaned data

#convert infection rate into integers
cleaned_data$infection_rate_10k_integer<-round(cleaned_data$infection_rate_10k,digits = 0)

#split data set into training and testing
#make this example reproducible
set.seed(1)

#use 70% of dataset as training set and 30% as test set
sample <- sample(c(TRUE, FALSE), nrow(cleaned_data), replace=TRUE, prob=c(0.8,0.2))
train  <- cleaned_data[sample, ]
test   <- cleaned_data[!sample, ]

#fit NB model to training set
summary(m1 <- glm.nb(infection_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pctwhite+pctblack+pcthispanic+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity,data = train))
vif(m1)


#RMSE for train data
#create new data set using independent varaibles
train_predict<- subset(train, select = -c(infection_rate_10k_integer))
#predict usin fitted NB 
train$predict_infection<- predict(m1, train_predict,type ="response")
rmse_train<-rmse(train$infection_rate_10k_integer,train$predict_infection)
rmse_train

#Normalized RMSE train
Normalized_rmse_train<- rmse_train/mean(train$infection_rate_10k_integer)
Normalized_rmse_train


#RMSE for test data
test_predict<- subset(test, select = -c(infection_rate_10k_integer))
test$predict_infection<- predict(m1, test_predict,type = "response")
rmse_test<- rmse(test$infection_rate_10k_integer,test$predict_infection)
rmse_test


#Normalized RMSE test
Normalized_rmse_test<- rmse_test/mean(test$infection_rate_10k_integer)
Normalized_rmse_test
```

#Phase 2 high infection model 4
```{r}
#Removing outliers
summary(cluster_high_inf_dataphase2$infection_rate_10k)
iqr<-IQR(cluster_high_inf_dataphase2$infection_rate_10k)
Tmax = 1006.8+3*iqr #consider extreme outlires
Tmin = 815.2-3*iqr #consider extreme outlires
#outliers
outl<- subset(cluster_high_inf_dataphase2,cluster_high_inf_dataphase2$infection_rate_10k<Tmin| cluster_high_inf_dataphase2$infection_rate_10k>Tmax)
#cleaned data
cleaned_data<- subset(cluster_high_inf_dataphase2,cluster_high_inf_dataphase2$infection_rate_10k>Tmin & cluster_high_inf_dataphase2$infection_rate_10k<Tmax)
#normality test
shapiro.test(cleaned_data$infection_rate_10k)

#summary(cleaned_data$infection_rate_10k)
#Normalize data(since data is not normally distributed we cannot use z score)
#cleaned_data$normalized_infection_rate_10k<- (cleaned_data$infection_rate_10k-63.17)/(321.49-63.17)

#(But fitting NB model to a standardized data doesn't make any sense.Therefore we decided to fit a model after removing outliers)

#Now I fit the NB model for infection_rate_10k in cleaned data

#convert infection rate into integers
cleaned_data$infection_rate_10k_integer<-round(cleaned_data$infection_rate_10k,digits = 0)

#split data set into training and testing
#make this example reproducible
set.seed(1)

#use 70% of dataset as training set and 30% as test set
sample <- sample(c(TRUE, FALSE), nrow(cleaned_data), replace=TRUE, prob=c(0.8,0.2))
train  <- cleaned_data[sample, ]
test   <- cleaned_data[!sample, ]

#fit NB model to training set
summary(m1 <- glm.nb(infection_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pctwhite+pctblack+pcthispanic+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity,data = train))
vif(m1)

summary(m1 <- glm.nb(infection_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pctblack+pcthispanic+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity,data = train))
vif(m1)

#RMSE for train data
#create new data set using independent varaibles
train_predict<- subset(train, select = -c(infection_rate_10k_integer))
#predict usin fitted NB 
train$predict_infection<- predict(m1, train_predict,type ="response")
rmse_train<-rmse(train$infection_rate_10k_integer,train$predict_infection)
rmse_train

#Normalized RMSE train
Normalized_rmse_train<- rmse_train/mean(train$infection_rate_10k_integer)
Normalized_rmse_train


#RMSE for test data
test_predict<- subset(test, select = -c(infection_rate_10k_integer))
test$predict_infection<- predict(m1, test_predict,type = "response")
rmse_test<- rmse(test$infection_rate_10k_integer,test$predict_infection)
rmse_test


#Normalized RMSE test
Normalized_rmse_test<- rmse_test/mean(test$infection_rate_10k_integer)
Normalized_rmse_test
```


#Phase 2 medium infection model 5
```{r}
#Removing outliers
summary(cluster_medium_inf_dataphase2$infection_rate_10k)
iqr<-IQR(cluster_medium_inf_dataphase2$infection_rate_10k)
Tmax = 672.4+3*iqr #consider extreme outlires
Tmin = 526.0-3*iqr #consider extreme outlires
#outliers
outl<- subset(cluster_medium_inf_dataphase2,cluster_medium_inf_dataphase2$infection_rate_10k<Tmin| cluster_medium_inf_dataphase2$infection_rate_10k>Tmax)
#cleaned data
cleaned_data<- subset(cluster_medium_inf_dataphase2,cluster_medium_inf_dataphase2$infection_rate_10k>Tmin & cluster_medium_inf_dataphase2$infection_rate_10k<Tmax)
#normality test
shapiro.test(cleaned_data$infection_rate_10k)

#summary(cleaned_data$infection_rate_10k)
#Normalize data(since data is not normally distributed we cannot use z score)
#cleaned_data$normalized_infection_rate_10k<- (cleaned_data$infection_rate_10k-63.17)/(321.49-63.17)

#(But fitting NB model to a standardized data doesn't make any sense.Therefore we decided to fit a model after removing outliers)

#Now I fit the NB model for infection_rate_10k in cleaned data

#convert infection rate into integers
cleaned_data$infection_rate_10k_integer<-round(cleaned_data$infection_rate_10k,digits = 0)

#split data set into training and testing
#make this example reproducible
set.seed(1)

#use 70% of dataset as training set and 30% as test set
sample <- sample(c(TRUE, FALSE), nrow(cleaned_data), replace=TRUE, prob=c(0.8,0.2))
train  <- cleaned_data[sample, ]
test   <- cleaned_data[!sample, ]

#fit NB model to training set
summary(m1 <- glm.nb(infection_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pctwhite+pctblack+pcthispanic+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity,data = train))
vif(m1)

summary(m1 <- glm.nb(infection_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pctblack+pcthispanic+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity,data = train))
vif(m1)

#RMSE for train data
#create new data set using independent varaibles
train_predict<- subset(train, select = -c(infection_rate_10k_integer))
#predict usin fitted NB 
train$predict_infection<- predict(m1, train_predict,type ="response")
rmse_train<-rmse(train$infection_rate_10k_integer,train$predict_infection)
rmse_train

#Normalized RMSE train
Normalized_rmse_train<- rmse_train/mean(train$infection_rate_10k_integer)
Normalized_rmse_train


#RMSE for test data
test_predict<- subset(test, select = -c(infection_rate_10k_integer))
test$predict_infection<- predict(m1, test_predict,type = "response")
rmse_test<- rmse(test$infection_rate_10k_integer,test$predict_infection)
rmse_test


#Normalized RMSE test
Normalized_rmse_test<- rmse_test/mean(test$infection_rate_10k_integer)
Normalized_rmse_test
```


#Phase 2 low infection model 6
```{r}
#Removing outliers
summary(cluster_low_inf_dataphase2$infection_rate_10k)
iqr<-IQR(cluster_low_inf_dataphase2$infection_rate_10k)
Tmax = 395.4+3*iqr #consider extreme outlires
Tmin = 250.1-3*iqr #consider extreme outlires
#outliers
outl<- subset(cluster_low_inf_dataphase2,cluster_low_inf_dataphase2$infection_rate_10k<Tmin| cluster_low_inf_dataphase2$infection_rate_10k>Tmax)
#cleaned data
cleaned_data<- subset(cluster_low_inf_dataphase2,cluster_low_inf_dataphase2$infection_rate_10k>Tmin & cluster_low_inf_dataphase2$infection_rate_10k<Tmax)
#normality test
shapiro.test(cleaned_data$infection_rate_10k)

#summary(cleaned_data$infection_rate_10k)
#Normalize data(since data is not normally distributed we cannot use z score)
#cleaned_data$normalized_infection_rate_10k<- (cleaned_data$infection_rate_10k-63.17)/(321.49-63.17)

#(But fitting NB model to a standardized data doesn't make any sense.Therefore we decided to fit a model after removing outliers)

#Now I fit the NB model for infection_rate_10k in cleaned data

#convert infection rate into integers
cleaned_data$infection_rate_10k_integer<-round(cleaned_data$infection_rate_10k,digits = 0)

#split data set into training and testing
#make this example reproducible
set.seed(1)

#use 70% of dataset as training set and 30% as test set
sample <- sample(c(TRUE, FALSE), nrow(cleaned_data), replace=TRUE, prob=c(0.8,0.2))
train  <- cleaned_data[sample, ]
test   <- cleaned_data[!sample, ]

#fit NB model to training set
summary(m1 <- glm.nb(infection_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pctwhite+pctblack+pcthispanic+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity,data = train))
vif(m1)

summary(m1 <- glm.nb(infection_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pctblack+pcthispanic+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity,data = train))
vif(m1)

#RMSE for train data
#create new data set using independent varaibles
train_predict<- subset(train, select = -c(infection_rate_10k_integer))
#predict usin fitted NB 
train$predict_infection<- predict(m1, train_predict,type ="response")
rmse_train<-rmse(train$infection_rate_10k_integer,train$predict_infection)
rmse_train

#Normalized RMSE train
Normalized_rmse_train<- rmse_train/mean(train$infection_rate_10k_integer)
Normalized_rmse_train


#RMSE for test data
test_predict<- subset(test, select = -c(infection_rate_10k_integer))
test$predict_infection<- predict(m1, test_predict,type = "response")
rmse_test<- rmse(test$infection_rate_10k_integer,test$predict_infection)
rmse_test


#Normalized RMSE test
Normalized_rmse_test<- rmse_test/mean(test$infection_rate_10k_integer)
Normalized_rmse_test
```

#All year high infection model 7
```{r}
#Removing outliers
summary(cluster_high_inf_datafull$infection_rate_10k)
iqr<-IQR(cluster_high_inf_datafull$infection_rate_10k)
Tmax = 1082.3+3*iqr #consider extreme outlires
Tmin = 876.1 -3*iqr #consider extreme outlires
#outliers
outl<- subset(cluster_high_inf_datafull,cluster_high_inf_datafull$infection_rate_10k<Tmin| cluster_high_inf_datafull$infection_rate_10k>Tmax)
#cleaned data
cleaned_data<- subset(cluster_high_inf_datafull,cluster_high_inf_datafull$infection_rate_10k>Tmin & cluster_high_inf_datafull$infection_rate_10k<Tmax)
#normality test
shapiro.test(cleaned_data$infection_rate_10k)

#summary(cleaned_data$infection_rate_10k)
#Normalize data(since data is not normally distributed we cannot use z score)
#cleaned_data$normalized_infection_rate_10k<- (cleaned_data$infection_rate_10k-63.17)/(321.49-63.17)

#(But fitting NB model to a standardized data doesn't make any sense.Therefore we decided to fit a model after removing outliers)

#Now I fit the NB model for infection_rate_10k in cleaned data

#convert infection rate into integers
cleaned_data$infection_rate_10k_integer<-round(cleaned_data$infection_rate_10k,digits = 0)

#split data set into training and testing
#make this example reproducible
set.seed(1)

#use 70% of dataset as training set and 30% as test set
sample <- sample(c(TRUE, FALSE), nrow(cleaned_data), replace=TRUE, prob=c(0.8,0.2))
train  <- cleaned_data[sample, ]
test   <- cleaned_data[!sample, ]

#fit NB model to training set
summary(m1 <- glm.nb(infection_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pctwhite+pctblack+pcthispanic+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity,data = train))
vif(m1)

summary(m1 <- glm.nb(infection_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pctblack+pcthispanic+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity,data = train))
vif(m1)


#RMSE for train data
#create new data set using independent varaibles
train_predict<- subset(train, select = -c(infection_rate_10k_integer))
#predict usin fitted NB 
train$predict_infection<- predict(m1, train_predict,type ="response")
rmse_train<-rmse(train$infection_rate_10k_integer,train$predict_infection)
rmse_train

#Normalized RMSE train
Normalized_rmse_train<- rmse_train/mean(train$infection_rate_10k_integer)
Normalized_rmse_train


#RMSE for test data
test_predict<- subset(test, select = -c(infection_rate_10k_integer))
test$predict_infection<- predict(m1, test_predict,type = "response")
rmse_test<- rmse(test$infection_rate_10k_integer,test$predict_infection)
rmse_test


#Normalized RMSE test
Normalized_rmse_test<- rmse_test/mean(test$infection_rate_10k_integer)
Normalized_rmse_test
```

#All year medium infection model 8
```{r}
#Removing outliers
summary(cluster_medium_inf_datafull$infection_rate_10k)
iqr<-IQR(cluster_medium_inf_datafull$infection_rate_10k)
Tmax = 724.5+3*iqr #consider extreme outlires
Tmin = 577.1 -3*iqr #consider extreme outlires
#outliers
outl<- subset(cluster_medium_inf_datafull,cluster_medium_inf_datafull$infection_rate_10k<Tmin| cluster_medium_inf_datafull$infection_rate_10k>Tmax)
#cleaned data
cleaned_data<- subset(cluster_medium_inf_datafull,cluster_medium_inf_datafull$infection_rate_10k>Tmin & cluster_medium_inf_datafull$infection_rate_10k<Tmax)
#normality test
shapiro.test(cleaned_data$infection_rate_10k)

#summary(cleaned_data$infection_rate_10k)
#Normalize data(since data is not normally distributed we cannot use z score)
#cleaned_data$normalized_infection_rate_10k<- (cleaned_data$infection_rate_10k-63.17)/(321.49-63.17)

#(But fitting NB model to a standardized data doesn't make any sense.Therefore we decided to fit a model after removing outliers)

#Now I fit the NB model for infection_rate_10k in cleaned data

#convert infection rate into integers
cleaned_data$infection_rate_10k_integer<-round(cleaned_data$infection_rate_10k,digits = 0)

#split data set into training and testing
#make this example reproducible
set.seed(1)

#use 70% of dataset as training set and 30% as test set
sample <- sample(c(TRUE, FALSE), nrow(cleaned_data), replace=TRUE, prob=c(0.8,0.2))
train  <- cleaned_data[sample, ]
test   <- cleaned_data[!sample, ]

#fit NB model to training set
summary(m1 <- glm.nb(infection_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pctwhite+pctblack+pcthispanic+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity,data = train))
vif(m1)

summary(m1 <- glm.nb(infection_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pctblack+pcthispanic+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity,data = train))
vif(m1)


#RMSE for train data
#create new data set using independent varaibles
train_predict<- subset(train, select = -c(infection_rate_10k_integer))
#predict usin fitted NB 
train$predict_infection<- predict(m1, train_predict,type ="response")
rmse_train<-rmse(train$infection_rate_10k_integer,train$predict_infection)
rmse_train

#Normalized RMSE train
Normalized_rmse_train<- rmse_train/mean(train$infection_rate_10k_integer)
Normalized_rmse_train


#RMSE for test data
test_predict<- subset(test, select = -c(infection_rate_10k_integer))
test$predict_infection<- predict(m1, test_predict,type = "response")
rmse_test<- rmse(test$infection_rate_10k_integer,test$predict_infection)
rmse_test


#Normalized RMSE test
Normalized_rmse_test<- rmse_test/mean(test$infection_rate_10k_integer)
Normalized_rmse_test
```

#All year low infection model 9
```{r}
#Removing outliers
summary(cluster_low_inf_datafull$infection_rate_10k)
iqr<-IQR(cluster_low_inf_datafull$infection_rate_10k)
Tmax = 439.25+3*iqr #consider extreme outlires
Tmin = 281.90 -3*iqr #consider extreme outlires
#outliers
outl<- subset(cluster_low_inf_datafull,cluster_low_inf_datafull$infection_rate_10k<Tmin| cluster_low_inf_datafull$infection_rate_10k>Tmax)
#cleaned data
cleaned_data<- subset(cluster_low_inf_datafull,cluster_low_inf_datafull$infection_rate_10k>Tmin & cluster_low_inf_datafull$infection_rate_10k<Tmax)
#normality test
shapiro.test(cleaned_data$infection_rate_10k)

#summary(cleaned_data$infection_rate_10k)
#Normalize data(since data is not normally distributed we cannot use z score)
#cleaned_data$normalized_infection_rate_10k<- (cleaned_data$infection_rate_10k-63.17)/(321.49-63.17)

#(But fitting NB model to a standardized data doesn't make any sense.Therefore we decided to fit a model after removing outliers)

#Now I fit the NB model for infection_rate_10k in cleaned data

#convert infection rate into integers
cleaned_data$infection_rate_10k_integer<-round(cleaned_data$infection_rate_10k,digits = 0)

#split data set into training and testing
#make this example reproducible
set.seed(1)

#use 70% of dataset as training set and 30% as test set
sample <- sample(c(TRUE, FALSE), nrow(cleaned_data), replace=TRUE, prob=c(0.8,0.2))
train  <- cleaned_data[sample, ]
test   <- cleaned_data[!sample, ]

#fit NB model to training set
summary(m1 <- glm.nb(infection_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pctwhite+pctblack+pcthispanic+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity,data = train))
vif(m1)

summary(m1 <- glm.nb(infection_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pctblack+pcthispanic+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity,data = train))
vif(m1)


#RMSE for train data
#create new data set using independent varaibles
train_predict<- subset(train, select = -c(infection_rate_10k_integer))
#predict usin fitted NB 
train$predict_infection<- predict(m1, train_predict,type ="response")
rmse_train<-rmse(train$infection_rate_10k_integer,train$predict_infection)
rmse_train

#Normalized RMSE train
Normalized_rmse_train<- rmse_train/mean(train$infection_rate_10k_integer)
Normalized_rmse_train


#RMSE for test data
test_predict<- subset(test, select = -c(infection_rate_10k_integer))
test$predict_infection<- predict(m1, test_predict,type = "response")
rmse_test<- rmse(test$infection_rate_10k_integer,test$predict_infection)
rmse_test


#Normalized RMSE test
Normalized_rmse_test<- rmse_test/mean(test$infection_rate_10k_integer)
Normalized_rmse_test
```

##############################################
Interaction effect for Infection rate clusters
##############################################


#Phase 2 medium infection model 5(This one found as a good models using NRMSE)
```{r}
#Removing outliers
summary(cluster_medium_inf_dataphase2$infection_rate_10k)
iqr<-IQR(cluster_medium_inf_dataphase2$infection_rate_10k)
Tmax = 672.4+3*iqr #consider extreme outlires
Tmin = 526.0-3*iqr #consider extreme outlires
#outliers
outl<- subset(cluster_medium_inf_dataphase2,cluster_medium_inf_dataphase2$infection_rate_10k<Tmin| cluster_medium_inf_dataphase2$infection_rate_10k>Tmax)
#cleaned data
cleaned_data<- subset(cluster_medium_inf_dataphase2,cluster_medium_inf_dataphase2$infection_rate_10k>Tmin & cluster_medium_inf_dataphase2$infection_rate_10k<Tmax)
#normality test
shapiro.test(cleaned_data$infection_rate_10k)

#summary(cleaned_data$infection_rate_10k)
#Normalize data(since data is not normally distributed we cannot use z score)
#cleaned_data$normalized_infection_rate_10k<- (cleaned_data$infection_rate_10k-63.17)/(321.49-63.17)

#(But fitting NB model to a standardized data doesn't make any sense.Therefore we decided to fit a model after removing outliers)

#Now I fit the NB model for infection_rate_10k in cleaned data

#convert infection rate into integers
cleaned_data$infection_rate_10k_integer<-round(cleaned_data$infection_rate_10k,digits = 0)

#split data set into training and testing
#make this example reproducible
set.seed(1)

#use 80% of dataset as training set and 20% as test set
sample <- sample(c(TRUE, FALSE), nrow(cleaned_data), replace=TRUE, prob=c(0.8,0.2))
train  <- cleaned_data[sample, ]
test   <- cleaned_data[!sample, ]

#fit NB model to training set
summary(m1 <- glm.nb(infection_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pctwhite+pctblack+pcthispanic+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity,data = train))
vif(m1)

summary(m1 <- glm.nb(infection_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pctblack+pcthispanic+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity,data = train))
vif(m1)

#model only considering significant varaiables
summary(m1 <- glm.nb(infection_rate_10k_integer~popdense+ageab65+pctpov,data = train))
vif(m1)

#with Interaction effect(Uncentered)
summary(m1 <- glm.nb(infection_rate_10k_integer~popdense+ageab65+pctblack+popdense*ageab65+popdense*pctblack+ageab65*pctblack,data = train))
#multi-collinearity
vif(m1)


#with Interaction effect(Uncentered)- consider all the varaibles for the analysis
summary(m1 <- glm.nb(infection_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity + pctblack+pcthispanic + 
                       popdense*ageab65 + popdense*pctpov + popdense*pctmale+ popdense*pct_highschool_or_less + popdense * pct_Adults_with_Diabetes + popdense* pct_Adults_with_Obesity + 
                       ageab65*pctpov+ ageab65*pctmale +ageab65* pct_highschool_or_less+ ageab65* pct_Adults_with_Diabetes + ageab65 *pct_Adults_with_Obesity + 
                       pctpov*pctmale + pctpov*pct_highschool_or_less + pctpov *pct_Adults_with_Diabetes + pctpov*pct_Adults_with_Obesity + 
                       pctmale* pct_highschool_or_less + pctmale * pct_Adults_with_Diabetes + pctmale* pct_Adults_with_Obesity + 
                       pct_highschool_or_less * pct_Adults_with_Diabetes + pct_highschool_or_less * pct_Adults_with_Obesity + 
                       pct_Adults_with_Diabetes * pct_Adults_with_Obesity +
                       
                       pctblack*popdense+pctblack*ageab65+pctblack*pctpov+pctblack*pctmale+pctblack*pct_highschool_or_less+pctblack*pct_Adults_with_Diabetes+pctblack*pct_Adults_with_Obesity +
                       
                       pcthispanic*popdense+pcthispanic*ageab65+pcthispanic*pctpov+pcthispanic*pctmale+pcthispanic*pct_highschool_or_less+pcthispanic*pct_Adults_with_Diabetes+pcthispanic*pct_Adults_with_Obesity + pcthispanic*pctblack,data = train))

#RMSE for train data
#create new data set using independent varaibles
train_predict<- subset(train, select = -c(infection_rate_10k_integer))
#predict usin fitted NB 
train$predict_infection<- predict(m1, train_predict,type ="response")
rmse_train<-rmse(train$infection_rate_10k_integer,train$predict_infection)
rmse_train

#Normalized RMSE train
Normalized_rmse_train<- rmse_train/mean(train$infection_rate_10k_integer)
Normalized_rmse_train


#RMSE for test data
test_predict<- subset(test, select = -c(infection_rate_10k_integer))
test$predict_infection<- predict(m1, test_predict,type = "response")
rmse_test<- rmse(test$infection_rate_10k_integer,test$predict_infection)
rmse_test


#Normalized RMSE test
Normalized_rmse_test<- rmse_test/mean(test$infection_rate_10k_integer)
Normalized_rmse_test


#RMSE for train data
#create new data set using independent varaibles
train_predict<- subset(train, select = -c(infection_rate_10k_integer))
#predict usin fitted NB 
train$predict_infection<- predict(m1, train_predict,type ="response")
rmse_train<-rmse(train$infection_rate_10k_integer,train$predict_infection)
rmse_train

#Normalized RMSE train
Normalized_rmse_train<- rmse_train/mean(train$infection_rate_10k_integer)
Normalized_rmse_train


#RMSE for test data
test_predict<- subset(test, select = -c(infection_rate_10k_integer))
test$predict_infection<- predict(m1, test_predict,type = "response")
rmse_test<- rmse(test$infection_rate_10k_integer,test$predict_infection)
rmse_test


#Normalized RMSE test
Normalized_rmse_test<- rmse_test/mean(test$infection_rate_10k_integer)
Normalized_rmse_test

#Furthermore, try to improve model considering variables that are not actually significant in the model but close to be significnat.

#original model again
summary(m1 <- glm.nb(infection_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity,data = train))
vif(m1)
#in this case there is no such a variable


#Phase 2 high infection model 4(This one found as a good models using NRMSE)
```{r}
#Removing outliers
summary(cluster_high_inf_dataphase2$infection_rate_10k)
iqr<-IQR(cluster_high_inf_dataphase2$infection_rate_10k)
Tmax = 1006.8+3*iqr #consider extreme outlires
Tmin = 815.2-3*iqr #consider extreme outlires
#outliers
outl<- subset(cluster_high_inf_dataphase2,cluster_high_inf_dataphase2$infection_rate_10k<Tmin| cluster_high_inf_dataphase2$infection_rate_10k>Tmax)
#cleaned data
cleaned_data<- subset(cluster_high_inf_dataphase2,cluster_high_inf_dataphase2$infection_rate_10k>Tmin & cluster_high_inf_dataphase2$infection_rate_10k<Tmax)
#normality test
shapiro.test(cleaned_data$infection_rate_10k)

#summary(cleaned_data$infection_rate_10k)
#Normalize data(since data is not normally distributed we cannot use z score)
#cleaned_data$normalized_infection_rate_10k<- (cleaned_data$infection_rate_10k-63.17)/(321.49-63.17)

#(But fitting NB model to a standardized data doesn't make any sense.Therefore we decided to fit a model after removing outliers)

#Now I fit the NB model for infection_rate_10k in cleaned data

#convert infection rate into integers
cleaned_data$infection_rate_10k_integer<-round(cleaned_data$infection_rate_10k,digits = 0)

#split data set into training and testing
#make this example reproducible
set.seed(1)

#use 70% of dataset as training set and 30% as test set
sample <- sample(c(TRUE, FALSE), nrow(cleaned_data), replace=TRUE, prob=c(0.8,0.2))
train  <- cleaned_data[sample, ]
test   <- cleaned_data[!sample, ]

#fit NB model to training set
summary(m1 <- glm.nb(infection_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pctwhite+pctblack+pcthispanic+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity,data = train))
vif(m1)

summary(m1 <- glm.nb(infection_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pctblack+pcthispanic+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity,data = train))
vif(m1)



#with Interaction effect(Uncentered)- consider all the varaibles for the analysis
summary(m1 <- glm.nb(infection_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity + pctblack + popdense*ageab65 + popdense*pctpov + popdense*pctmale+ popdense*pct_highschool_or_less + popdense * pct_Adults_with_Diabetes + popdense* pct_Adults_with_Obesity + ageab65*pctpov+ ageab65*pctmale +ageab65* pct_highschool_or_less+ ageab65* pct_Adults_with_Diabetes + ageab65 *pct_Adults_with_Obesity + pctpov*pctmale + pctpov*pct_highschool_or_less + pctpov *pct_Adults_with_Diabetes + pctpov*pct_Adults_with_Obesity + pctmale* pct_highschool_or_less + pctmale * pct_Adults_with_Diabetes + pctmale* pct_Adults_with_Obesity + pct_highschool_or_less * pct_Adults_with_Diabetes + pct_highschool_or_less * pct_Adults_with_Obesity + pct_Adults_with_Diabetes * pct_Adults_with_Obesity +
                       pctblack*popdense+pctblack*ageab65+pctblack*pctpov+pctblack*pctmale+pctblack*pct_highschool_or_less+pctblack*pct_Adults_with_Diabetes+pctblack*pct_Adults_with_Obesity +
                       pctwhite*popdense+pctwhite*ageab65+pctwhite*pctpov+pctwhite*pctmale+pctwhite*pct_highschool_or_less+pctwhite*pct_Adults_with_Diabetes+pctwhite*pct_Adults_with_Obesity + pctwhite*pctblack+
                       pcthispanic*popdense+pcthispanic*ageab65+pcthispanic*pctpov+pcthispanic*pctmale+pcthispanic*pct_highschool_or_less+pcthispanic*pct_Adults_with_Diabetes+pcthispanic*pct_Adults_with_Obesity + pcthispanic*pctblack,data = train))



#RMSE for train data
#create new data set using independent varaibles
train_predict<- subset(train, select = -c(infection_rate_10k_integer))
#predict usin fitted NB 
train$predict_infection<- predict(m1, train_predict,type ="response")
rmse_train<-rmse(train$infection_rate_10k_integer,train$predict_infection)
rmse_train

#Normalized RMSE train
Normalized_rmse_train<- rmse_train/mean(train$infection_rate_10k_integer)
Normalized_rmse_train


#RMSE for test data
test_predict<- subset(test, select = -c(infection_rate_10k_integer))
test$predict_infection<- predict(m1, test_predict,type = "response")
rmse_test<- rmse(test$infection_rate_10k_integer,test$predict_infection)
rmse_test


#Normalized RMSE test
Normalized_rmse_test<- rmse_test/mean(test$infection_rate_10k_integer)
Normalized_rmse_test



############################
Regression Analysis
############################

FATALITY

#Phase 1 low fatality model 1(This one has all zeros for falaity.Thus no point of modeling)

#Phase 1 medium fatality model 2
```{r}
#Removing outliers

cluster_medium_fat_dataphase1<- filter_if(cluster_medium_fat_dataphase1, is.numeric, all_vars((.) != 0))
summary(cluster_medium_fat_dataphase1$fatality_rate_10k)
iqr<-IQR(cluster_medium_fat_dataphase1$fatality_rate_10k)
Tmax = 281.690+3*iqr #consider extreme outlires
Tmin = 114.286 -3*iqr #consider extreme outlires
#outliers
outl<- subset(cluster_medium_fat_dataphase1,cluster_medium_fat_dataphase1$fatality_rate_10k<Tmin| cluster_medium_fat_dataphase1$fatality_rate_10k>Tmax)
#cleaned data
cleaned_data<- subset(cluster_medium_fat_dataphase1,cluster_medium_fat_dataphase1$fatality_rate_10k>Tmin & cluster_medium_fat_dataphase1$fatality_rate_10k<Tmax)
#normality test
shapiro.test(cleaned_data$fatality_rate_10k)

#Now I fit the NB model for infection_rate_10k in cleaned data

#convert infection rate into integers
cleaned_data$fatality_rate_10k_integer<-round(cleaned_data$fatality_rate_10k,digits = 0)

#split data set into training and testing
#make this example reproducible
set.seed(1)

#use 70% of dataset as training set and 30% as test set
sample <- sample(c(TRUE, FALSE), nrow(cleaned_data), replace=TRUE, prob=c(0.8,0.2))
train  <- cleaned_data[sample, ]
test   <- cleaned_data[!sample, ]

#fit NB model to training set
summary(m1 <- glm.nb(fatality_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pctwhite+pctblack+pcthispanic+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity,data = train))
vif(m1)

summary(m1 <- glm.nb(fatality_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pctblack+pcthispanic+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity,data = train))
vif(m1)

#RMSE for train data
#create new data set using independent varaibles
train_predict<- subset(train, select = -c(fatality_rate_10k_integer))
#predict using fitted NB 
train$predict_fatality<- predict(m1, train_predict,type ="response")
rmse_train<-rmse(train$fatality_rate_10k_integer,train$predict_fatality)
rmse_train

#Normalized RMSE train
Normalized_rmse_train<- rmse_train/mean(train$fatality_rate_10k_integer)
Normalized_rmse_train


#RMSE for test data
test_predict<- subset(test, select = -c(fatality_rate_10k_integer))
test$predict_fatality<- predict(m1, test_predict,type = "response")
rmse_test<- rmse(test$fatality_rate_10k_integer,test$predict_fatality)
rmse_test


#Normalized RMSE test
Normalized_rmse_test<- rmse_test/mean(test$fatality_rate_10k_integer)
Normalized_rmse_test
```

#Phase 1 high fatality model 3

```{r}
#Removing outliers

summary(cluster_high_fat_dataphase1$fatality_rate_10k)
iqr<-IQR(cluster_high_fat_dataphase1$fatality_rate_10k)
Tmax = 1000.0+3*iqr #consider extreme outlires
Tmin = 553.6 -3*iqr #consider extreme outlires
#outliers
outl<- subset(cluster_high_fat_dataphase1,cluster_high_fat_dataphase1$fatality_rate_10k<Tmin| cluster_high_fat_dataphase1$fatality_rate_10k>Tmax)
#cleaned data
cleaned_data<- subset(cluster_high_fat_dataphase1,cluster_high_fat_dataphase1$fatality_rate_10k>Tmin & cluster_high_fat_dataphase1$fatality_rate_10k<Tmax)
#normality test
shapiro.test(cleaned_data$fatality_rate_10k)

#Now I fit the NB model for infection_rate_10k in cleaned data

#convert infection rate into integers
cleaned_data$fatality_rate_10k_integer<-round(cleaned_data$fatality_rate_10k,digits = 0)

#split data set into training and testing
#make this example reproducible
set.seed(1)

#use 70% of dataset as training set and 30% as test set
sample <- sample(c(TRUE, FALSE), nrow(cleaned_data), replace=TRUE, prob=c(0.8,0.2))
train  <- cleaned_data[sample, ]
test   <- cleaned_data[!sample, ]

#fit NB model to training set
summary(m1 <- glm.nb(fatality_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pctwhite+pctblack+pcthispanic+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity,data = train))
vif(m1)

summary(m1 <- glm.nb(fatality_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pctblack+pcthispanic+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity,data = train))
vif(m1)

#RMSE for train data
#create new data set using independent varaibles
train_predict<- subset(train, select = -c(fatality_rate_10k_integer))
#predict usin fitted NB 
train$predict_fatality<- predict(m1, train_predict,type ="response")
rmse_train<-rmse(train$fatality_rate_10k_integer,train$predict_fatality)
rmse_train

#Normalized RMSE train
Normalized_rmse_train<- rmse_train/mean(train$fatality_rate_10k_integer)
Normalized_rmse_train


#RMSE for test data
test_predict<- subset(test, select = -c(fatality_rate_10k_integer))
test$predict_fatality<- predict(m1, test_predict,type = "response")
rmse_test<- rmse(test$fatality_rate_10k_integer,test$predict_fatality)
rmse_test


#Normalized RMSE test
Normalized_rmse_test<- rmse_test/mean(test$fatality_rate_10k_integer)
Normalized_rmse_test
```



#Phase 2 low fatality model 4

```{r}
#Removing outliers

cluster_low_fat_dataphase2<- filter_if(cluster_low_fat_dataphase2, is.numeric, all_vars((.) != 0))
summary(cluster_low_fat_dataphase2$fatality_rate_10k)
iqr<-IQR(cluster_low_fat_dataphase2$fatality_rate_10k)
Tmax = 78.302+3*iqr #consider extreme outlires
Tmin = 47.111 -3*iqr #consider extreme outlires
#outliers
outl<- subset(cluster_low_fat_dataphase2,cluster_low_fat_dataphase2$fatality_rate_10k<Tmin| cluster_low_fat_dataphase2$fatality_rate_10k>Tmax)
#cleaned data
cleaned_data<- subset(cluster_low_fat_dataphase2,cluster_low_fat_dataphase2$fatality_rate_10k>Tmin & cluster_low_fat_dataphase2$fatality_rate_10k<Tmax)
#normality test
shapiro.test(cleaned_data$fatality_rate_10k)

#Now I fit the NB model for infection_rate_10k in cleaned data

#convert infection rate into integers
cleaned_data$fatality_rate_10k_integer<-round(cleaned_data$fatality_rate_10k,digits = 0)

#split data set into training and testing
#make this example reproducible
set.seed(1)

#use 70% of dataset as training set and 30% as test set
sample <- sample(c(TRUE, FALSE), nrow(cleaned_data), replace=TRUE, prob=c(0.8,0.2))
train  <- cleaned_data[sample, ]
test   <- cleaned_data[!sample, ]

#fit NB model to training set
summary(m1 <- glm.nb(fatality_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pctwhite+pctblack+pcthispanic+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity,data = train))
vif(m1)


#RMSE for train data
#create new data set using independent varaibles
train_predict<- subset(train, select = -c(fatality_rate_10k_integer))
#predict usin fitted NB 
train$predict_fatality<- predict(m1, train_predict,type ="response")
rmse_train<-rmse(train$fatality_rate_10k_integer,train$predict_fatality)
rmse_train

#Normalized RMSE train
Normalized_rmse_train<- rmse_train/mean(train$fatality_rate_10k_integer)
Normalized_rmse_train


#RMSE for test data
test_predict<- subset(test, select = -c(fatality_rate_10k_integer))
test$predict_fatality<- predict(m1, test_predict,type = "response")
rmse_test<- rmse(test$fatality_rate_10k_integer,test$predict_fatality)
rmse_test


#Normalized RMSE test
Normalized_rmse_test<- rmse_test/mean(test$fatality_rate_10k_integer)
Normalized_rmse_test
```


#Phase 2 medium fatality model 5

```{r}
#Removing outliers

summary(cluster_medium_fat_dataphase2$fatality_rate_10k)
iqr<-IQR(cluster_medium_fat_dataphase2$fatality_rate_10k)
Tmax = 165.80+3*iqr #consider extreme outlires
Tmin = 112.35 -3*iqr #consider extreme outlires
#outliers
outl<- subset(cluster_medium_fat_dataphase2,cluster_medium_fat_dataphase2$fatality_rate_10k<Tmin| cluster_medium_fat_dataphase2$fatality_rate_10k>Tmax)
#cleaned data
cleaned_data<- subset(cluster_medium_fat_dataphase2,cluster_medium_fat_dataphase2$fatality_rate_10k>Tmin & cluster_medium_fat_dataphase2$fatality_rate_10k<Tmax)
#normality test
shapiro.test(cleaned_data$fatality_rate_10k)

#Now I fit the NB model for infection_rate_10k in cleaned data

#convert infection rate into integers
cleaned_data$fatality_rate_10k_integer<-round(cleaned_data$fatality_rate_10k,digits = 0)

#split data set into training and testing
#make this example reproducible
set.seed(1)

#use 70% of dataset as training set and 30% as test set
sample <- sample(c(TRUE, FALSE), nrow(cleaned_data), replace=TRUE, prob=c(0.8,0.2))
train  <- cleaned_data[sample, ]
test   <- cleaned_data[!sample, ]

#fit NB model to training set
summary(m1 <- glm.nb(fatality_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pctwhite+pctblack+pcthispanic+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity,data = train))
vif(m1)

summary(m1 <- glm.nb(fatality_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pctblack+pcthispanic+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity,data = train))
vif(m1)

#RMSE for train data
#create new data set using independent varaibles
train_predict<- subset(train, select = -c(fatality_rate_10k_integer))
#predict usin fitted NB 
train$predict_fatality<- predict(m1, train_predict,type ="response")
rmse_train<-rmse(train$fatality_rate_10k_integer,train$predict_fatality)
rmse_train

#Normalized RMSE train
Normalized_rmse_train<- rmse_train/mean(train$fatality_rate_10k_integer)
Normalized_rmse_train


#RMSE for test data
test_predict<- subset(test, select = -c(fatality_rate_10k_integer))
test$predict_fatality<- predict(m1, test_predict,type = "response")
rmse_test<- rmse(test$fatality_rate_10k_integer,test$predict_fatality)
rmse_test


#Normalized RMSE test
Normalized_rmse_test<- rmse_test/mean(test$fatality_rate_10k_integer)
Normalized_rmse_test
```


#Phase 2 high fatality model 6

```{r}
#Removing outliers

summary(cluster_high_fat_dataphase2$fatality_rate_10k)
iqr<-IQR(cluster_high_fat_dataphase2$fatality_rate_10k)
Tmax = 330.6+3*iqr #consider extreme outlires
Tmin = 232.6 -3*iqr #consider extreme outlires
#outliers
outl<- subset(cluster_high_fat_dataphase2,cluster_high_fat_dataphase2$fatality_rate_10k<Tmin| cluster_high_fat_dataphase2$fatality_rate_10k>Tmax)
#cleaned data
cleaned_data<- subset(cluster_high_fat_dataphase2,cluster_high_fat_dataphase2$fatality_rate_10k>Tmin & cluster_high_fat_dataphase2$fatality_rate_10k<Tmax)
#normality test
shapiro.test(cleaned_data$fatality_rate_10k)

#Now I fit the NB model for infection_rate_10k in cleaned data

#convert infection rate into integers
cleaned_data$fatality_rate_10k_integer<-round(cleaned_data$fatality_rate_10k,digits = 0)

#split data set into training and testing
#make this example reproducible
set.seed(1)

#use 70% of dataset as training set and 30% as test set
sample <- sample(c(TRUE, FALSE), nrow(cleaned_data), replace=TRUE, prob=c(0.8,0.2))
train  <- cleaned_data[sample, ]
test   <- cleaned_data[!sample, ]

#fit NB model to training set
summary(m1 <- glm.nb(fatality_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pctwhite+pctblack+pcthispanic+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity,data = train))
vif(m1)

summary(m1 <- glm.nb(fatality_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pctblack+pcthispanic+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity,data = train))
vif(m1)

#RMSE for train data
#create new data set using independent varaibles
train_predict<- subset(train, select = -c(fatality_rate_10k_integer))
#predict usin fitted NB 
train$predict_fatality<- predict(m1, train_predict,type ="response")
rmse_train<-rmse(train$fatality_rate_10k_integer,train$predict_fatality)
rmse_train

#Normalized RMSE train
Normalized_rmse_train<- rmse_train/mean(train$fatality_rate_10k_integer)
Normalized_rmse_train


#RMSE for test data
test_predict<- subset(test, select = -c(fatality_rate_10k_integer))
test$predict_fatality<- predict(m1, test_predict,type = "response")
rmse_test<- rmse(test$fatality_rate_10k_integer,test$predict_fatality)
rmse_test


#Normalized RMSE test
Normalized_rmse_test<- rmse_test/mean(test$fatality_rate_10k_integer)
Normalized_rmse_test
```

#All year low fatality model 7

```{r}
#Removing outliers

cluster_low_fat_datafull<- filter_if(cluster_low_fat_datafull, is.numeric, all_vars((.) != 0))
summary(cluster_low_fat_datafull$fatality_rate_10k)
iqr<-IQR(cluster_low_fat_datafull$fatality_rate_10k)
Tmax = 88.374 +3*iqr #consider extreme outlires
Tmin = 52.888 -3*iqr #consider extreme outlires
#outliers
outl<- subset(cluster_low_fat_datafull,cluster_low_fat_datafull$fatality_rate_10k<Tmin| cluster_low_fat_datafull$fatality_rate_10k>Tmax)
#cleaned data
cleaned_data<- subset(cluster_low_fat_datafull,cluster_low_fat_datafull$fatality_rate_10k>Tmin & cluster_low_fat_datafull$fatality_rate_10k<Tmax)
#normality test
shapiro.test(cleaned_data$fatality_rate_10k)

#Now I fit the NB model for infection_rate_10k in cleaned data

#convert infection rate into integers
cleaned_data$fatality_rate_10k_integer<-round(cleaned_data$fatality_rate_10k,digits = 0)

#split data set into training and testing
#make this example reproducible
set.seed(1)

#use 70% of dataset as training set and 30% as test set
sample <- sample(c(TRUE, FALSE), nrow(cleaned_data), replace=TRUE, prob=c(0.8,0.2))
train  <- cleaned_data[sample, ]
test   <- cleaned_data[!sample, ]

#fit NB model to training set
summary(m1 <- glm.nb(fatality_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pctwhite+pctblack+pcthispanic+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity,data = train))
vif(m1)


#RMSE for train data
#create new data set using independent varaibles
train_predict<- subset(train, select = -c(fatality_rate_10k_integer))
#predict usin fitted NB 
train$predict_fatality<- predict(m1, train_predict,type ="response")
rmse_train<-rmse(train$fatality_rate_10k_integer,train$predict_fatality)
rmse_train

#Normalized RMSE train
Normalized_rmse_train<- rmse_train/mean(train$fatality_rate_10k_integer)
Normalized_rmse_train


#RMSE for test data
test_predict<- subset(test, select = -c(fatality_rate_10k_integer))
test$predict_fatality<- predict(m1, test_predict,type = "response")
rmse_test<- rmse(test$fatality_rate_10k_integer,test$predict_fatality)
rmse_test


#Normalized RMSE test
Normalized_rmse_test<- rmse_test/mean(test$fatality_rate_10k_integer)
Normalized_rmse_test
```

#All year medium fatality model 8

```{r}
#Removing outliers

summary(cluster_medium_fat_datafull$fatality_rate_10k)
iqr<-IQR(cluster_medium_fat_datafull$fatality_rate_10k)
Tmax = 187.3 +3*iqr #consider extreme outlires
Tmin = 126.0  -3*iqr #consider extreme outlires
#outliers
outl<- subset(cluster_medium_fat_datafull,cluster_medium_fat_datafull$fatality_rate_10k<Tmin| cluster_medium_fat_datafull$fatality_rate_10k>Tmax)
#cleaned data
cleaned_data<- subset(cluster_medium_fat_datafull,cluster_medium_fat_datafull$fatality_rate_10k>Tmin & cluster_medium_fat_datafull$fatality_rate_10k<Tmax)
#normality test
shapiro.test(cleaned_data$fatality_rate_10k)

#Now I fit the NB model for infection_rate_10k in cleaned data

#convert infection rate into integers
cleaned_data$fatality_rate_10k_integer<-round(cleaned_data$fatality_rate_10k,digits = 0)

#split data set into training and testing
#make this example reproducible
set.seed(1)

#use 70% of dataset as training set and 30% as test set
sample <- sample(c(TRUE, FALSE), nrow(cleaned_data), replace=TRUE, prob=c(0.8,0.2))
train  <- cleaned_data[sample, ]
test   <- cleaned_data[!sample, ]

#fit NB model to training set
summary(m1 <- glm.nb(fatality_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pctwhite+pctblack+pcthispanic+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity,data = train))
vif(m1)

summary(m1 <- glm.nb(fatality_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pctblack+pcthispanic+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity,data = train))
vif(m1)


#RMSE for train data
#create new data set using independent varaibles
train_predict<- subset(train, select = -c(fatality_rate_10k_integer))
#predict usin fitted NB 
train$predict_fatality<- predict(m1, train_predict,type ="response")
rmse_train<-rmse(train$fatality_rate_10k_integer,train$predict_fatality)
rmse_train

#Normalized RMSE train
Normalized_rmse_train<- rmse_train/mean(train$fatality_rate_10k_integer)
Normalized_rmse_train


#RMSE for test data
test_predict<- subset(test, select = -c(fatality_rate_10k_integer))
test$predict_fatality<- predict(m1, test_predict,type = "response")
rmse_test<- rmse(test$fatality_rate_10k_integer,test$predict_fatality)
rmse_test


#Normalized RMSE test
Normalized_rmse_test<- rmse_test/mean(test$fatality_rate_10k_integer)
Normalized_rmse_test
```

#All year high fatality model 9

```{r}
#Removing outliers

summary(cluster_high_fat_datafull$fatality_rate_10k)
iqr<-IQR(cluster_high_fat_datafull$fatality_rate_10k)
Tmax = 356.7  +3*iqr #consider extreme outlires
Tmin = 256.6  -3*iqr #consider extreme outlires
#outliers
outl<- subset(cluster_high_fat_datafull,cluster_high_fat_datafull$fatality_rate_10k<Tmin| cluster_high_fat_datafull$fatality_rate_10k>Tmax)
#cleaned data
cleaned_data<- subset(cluster_high_fat_datafull,cluster_high_fat_datafull$fatality_rate_10k>Tmin & cluster_high_fat_datafull$fatality_rate_10k<Tmax)
#normality test
shapiro.test(cleaned_data$fatality_rate_10k)

#Now I fit the NB model for infection_rate_10k in cleaned data

#convert infection rate into integers
cleaned_data$fatality_rate_10k_integer<-round(cleaned_data$fatality_rate_10k,digits = 0)

#split data set into training and testing
#make this example reproducible
set.seed(1)

#use 70% of dataset as training set and 30% as test set
sample <- sample(c(TRUE, FALSE), nrow(cleaned_data), replace=TRUE, prob=c(0.8,0.2))
train  <- cleaned_data[sample, ]
test   <- cleaned_data[!sample, ]

#fit NB model to training set
summary(m1 <- glm.nb(fatality_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pctwhite+pctblack+pcthispanic+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity,data = train))
vif(m1)

summary(m1 <- glm.nb(fatality_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pctblack+pcthispanic+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity,data = train))
vif(m1)


#RMSE for train data
#create new data set using independent varaibles
train_predict<- subset(train, select = -c(fatality_rate_10k_integer))
#predict usin fitted NB 
train$predict_fatality<- predict(m1, train_predict,type ="response")
rmse_train<-rmse(train$fatality_rate_10k_integer,train$predict_fatality)
rmse_train

#Normalized RMSE train
Normalized_rmse_train<- rmse_train/mean(train$fatality_rate_10k_integer)
Normalized_rmse_train


#RMSE for test data
test_predict<- subset(test, select = -c(fatality_rate_10k_integer))
test$predict_fatality<- predict(m1, test_predict,type = "response")
rmse_test<- rmse(test$fatality_rate_10k_integer,test$predict_fatality)
rmse_test


#Normalized RMSE test
Normalized_rmse_test<- rmse_test/mean(test$fatality_rate_10k_integer)
Normalized_rmse_test
```

##########################
###Interaction effect - fatality###
##########################
#Phase 2 medium fatality model 5

```{r}
#Removing outliers

summary(cluster_medium_fat_dataphase2$fatality_rate_10k)
iqr<-IQR(cluster_medium_fat_dataphase2$fatality_rate_10k)
Tmax = 165.80+3*iqr #consider extreme outlires
Tmin = 112.35 -3*iqr #consider extreme outlires
#outliers
outl<- subset(cluster_medium_fat_dataphase2,cluster_medium_fat_dataphase2$fatality_rate_10k<Tmin| cluster_medium_fat_dataphase2$fatality_rate_10k>Tmax)
#cleaned data
cleaned_data<- subset(cluster_medium_fat_dataphase2,cluster_medium_fat_dataphase2$fatality_rate_10k>Tmin & cluster_medium_fat_dataphase2$fatality_rate_10k<Tmax)
#normality test
shapiro.test(cleaned_data$fatality_rate_10k)

#Now I fit the NB model for infection_rate_10k in cleaned data

#convert infection rate into integers
cleaned_data$fatality_rate_10k_integer<-round(cleaned_data$fatality_rate_10k,digits = 0)

#split data set into training and testing
#make this example reproducible
set.seed(1)

#use 70% of dataset as training set and 30% as test set
sample <- sample(c(TRUE, FALSE), nrow(cleaned_data), replace=TRUE, prob=c(0.8,0.2))
train  <- cleaned_data[sample, ]
test   <- cleaned_data[!sample, ]

#fit NB model to training set
summary(m1 <- glm.nb(fatality_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pctwhite+pctblack+pcthispanic+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity,data = train))
vif(m1)

summary(m1 <- glm.nb(fatality_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pctblack+pcthispanic+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity,data = train))
vif(m1)



#with Interaction effect(Uncentered)- consider all the varaibles for the analysis
summary(m1 <- glm.nb(fatality_rate_10k_integer~popdense+ageab65+pctpov+pctmale+pct_highschool_or_less+pct_Adults_with_Diabetes+pct_Adults_with_Obesity + pctblack+pcthispanic + 
                       popdense*ageab65 + popdense*pctpov + popdense*pctmale+ popdense*pct_highschool_or_less + popdense * pct_Adults_with_Diabetes + popdense* pct_Adults_with_Obesity + 
                       ageab65*pctpov+ ageab65*pctmale +ageab65* pct_highschool_or_less+ ageab65* pct_Adults_with_Diabetes + ageab65 *pct_Adults_with_Obesity + 
                       pctpov*pctmale + pctpov*pct_highschool_or_less + pctpov *pct_Adults_with_Diabetes + pctpov*pct_Adults_with_Obesity + 
                       pctmale* pct_highschool_or_less + pctmale * pct_Adults_with_Diabetes + pctmale* pct_Adults_with_Obesity + 
                       pct_highschool_or_less * pct_Adults_with_Diabetes + pct_highschool_or_less * pct_Adults_with_Obesity + 
                       pct_Adults_with_Diabetes * pct_Adults_with_Obesity +
                       
                       pctblack*popdense+pctblack*ageab65+pctblack*pctpov+pctblack*pctmale+pctblack*pct_highschool_or_less+pctblack*pct_Adults_with_Diabetes+pctblack*pct_Adults_with_Obesity +
                       
                       pcthispanic*popdense+pcthispanic*ageab65+pcthispanic*pctpov+pcthispanic*pctmale+pcthispanic*pct_highschool_or_less+pcthispanic*pct_Adults_with_Diabetes+pcthispanic*pct_Adults_with_Obesity + pcthispanic*pctblack,data = train))


#RMSE for train data
#create new data set using independent varaibles
train_predict<- subset(train, select = -c(fatality_rate_10k_integer))
#predict usin fitted NB 
train$predict_fatality<- predict(m1, train_predict,type ="response")
rmse_train<-rmse(train$fatality_rate_10k_integer,train$predict_fatality)
rmse_train

#Normalized RMSE train
Normalized_rmse_train<- rmse_train/mean(train$fatality_rate_10k_integer)
Normalized_rmse_train


#RMSE for test data
test_predict<- subset(test, select = -c(fatality_rate_10k_integer))
test$predict_fatality<- predict(m1, test_predict,type = "response")
rmse_test<- rmse(test$fatality_rate_10k_integer,test$predict_fatality)
rmse_test


#Normalized RMSE test
Normalized_rmse_test<- rmse_test/mean(test$fatality_rate_10k_integer)
Normalized_rmse_test



